Linear Vlasov Theory in the Shearing Sheet Approximation \\ with Application to the Magneto-Rotational Instability

### 2014-08-11 paper.tex
This branch is for submission to arXiv (v2)
Fixed typo
Added hybrid code references

The salient feature of having introduced $\vvc$ is that \cref{eq:orbit-eom-vc}
takes the same form as in a uniform plasma, with the gyro-frequency vector
$\vomg$ defined in \cref{eq:gyro-frequency-vector} playing the role of the
cyclotron frequency vector $\vomc=e\vec{B}/m$. Let us therefore adopt a
cylindrical coordinate system $(\vc_\parallel,\vc_\perp,\vartheta)$ in
$\vvc$-space that is aligned with $\vomg$. Here and in the following, $\perp$
and $\parallel$ will always be defined with respect to the direction $\vomg$.
The steady state Vlasov equation is
\begin{equation}
  [(\vec{v} + q\Omega x\ey)\times\vec{S}]\cdot\pder{f}{\vec{v}}
  = (\vvc\times\vomg)\cdot\pder{f}{\vvc} = -\omg\pder{f}{\vartheta} = 0,
\end{equation}
where the square of the gyration frequency $\omg$ is given by
\begin{equation}
  \label{eq:gyro-frequency}
  \omg^2 = S_y^2 + S_z(S_z - q\Omega).
\end{equation}
The equilibrium distribution function is thus gyrotropic in $\vvc$-space,
i.e.\ of the form $f(\vc_\perp,\vc_\parallel)$. For this reason we will refer
to $\vvc$ as the \emph{gyrotropic velocity}.

@@

The crude electron model represents the most severe limitation of Vlasov-fluid
theory. At the same time it may, however, also be viewed as the theory's
greatest merit. This is because the fast electron scales are not part of the
problem anymore: The plasma frequency and Debye length are not finite due to
charge-neutrality and neither is the electron Larmor motion due to the
vanishing electron mass. The absence of these fast time scales simplifies the
analysis and significantly reduces the computational burden.

>>>>

The salient feature of having introduced $\vvc$ is that \cref{eq:orbit-eom-vc}
takes the same form as in a uniform plasma, with the gyro-frequency vector
$\vomg$ defined in \cref{eq:gyro-frequency-vector} playing the role of the
cyclotron frequency vector $\vomc=e\vec{B}/m$. Let us therefore adopt a
cylindrical coordinate system $(\vc_\parallel,\vc_\perp,\vartheta)$ in
$\vvc$-space that is aligned with $\vomg$. Here and in the following, $\perp$
and $\parallel$ will always be defined with respect to the direction of
$\vomg$. The steady state Vlasov equation is
\begin{equation}
  [(\vec{v} + q\Omega x\ey)\times\vec{S}]\cdot\pder{f}{\vec{v}}
  = (\vvc\times\vomg)\cdot\pder{f}{\vvc} = -\omg\pder{f}{\vartheta} = 0,
\end{equation}
where the square of the gyration frequency $\omg$ is given by
\begin{equation}
  \label{eq:gyro-frequency}
  \omg^2 = S_y^2 + S_z(S_z - q\Omega).
\end{equation}
The equilibrium distribution function is thus gyrotropic in $\vvc$-space,
i.e.\ of the form $f(\vc_\perp,\vc_\parallel)$. For this reason we will refer
to $\vvc$ as the \emph{gyrotropic velocity}.

@@

The crude electron model represents the most severe limitation of Vlasov-fluid
theory. At the same time it may, however, also be viewed as the theory's
greatest merit. This is because the fast electron scales are not part of the
problem anymore: The plasma frequency and Debye length are not finite due to
charge-neutrality and neither is the electron Larmor motion due to the
vanishing electron mass. The absence of these fast time scales simplifies the
analysis and significantly reduces the computational burden in numerical
simulations \citep{Byers1978,Winske2003,Kunz2014}.
